---
title: "Quick start guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  color_pal<<-c("#E69F00",
              "#CC79A7","#56B4E9","#009E73",
              "#F0E442","#0072B2","#D55E00",
              "#E1BE6A","#0231A9"),
  myColor <- colorRampPalette(c("#FFFFFF",color_pal[4]))(50),
library(data.table),
library(pheatmap),
library(tidyverse),
library(magrittr)
)
```

```{r setup}
library(microSLAM)
n_tau=100 ## number of permutations for tau test
exp_metadata = read.csv("../example_data/exp_metadata.csv")
### read in example metadata
exp_genedata = read.csv("../example_data/genepresabs.csv") 
### read in example gene data
```


## Step 1 create GRM
### calculate the GRM (1-hamming distance between each sample) from gene data a samples by genes matrix
```{r}

GRM =calculate_GRM(exp_genedata)
```

### GRM visualization 

```{r pheatmap, echo=FALSE, fig.height=6, fig.width=8}
rownames(exp_metadata)=exp_metadata$sample_name
pheatmap(GRM,show_rownames=FALSE,show_colnames=FALSE,treeheight_row=0,treeheight_col = 0,labels_row="samples",labels_col="samples",
                    main=paste("GRM"),border_color=NA,annotation_row = exp_metadata[,-5],color=myColor,clustering_distance_rows=as.dist(1-GRM),clustering_distance_col=as.dist(1-GRM),clustering_method="average")
```

## Step 2 fit and test tau
### fit baseline glm for starting parameters in tau test
```{r}
glm_fit0=glm("y~age+1", data = exp_metadata, family = "binomial")
```

### fit tau test using baseline glm and GRM calculated above order of glm and GRM samples must match!!
```{r}
glmm_fit=fit_tau_test(glm_fit0, GRM,species_id="test",verbose = TRUE,log_file=NA)
```
### test the significance of the tau that was fit
```{r}
tautestfit=run_tau_test(glm_fit0, GRM,n_tau,species_id = "test", tau0=1, phi0=1)
```

### calculate the pvalue from the permutation test ran on tau
```{r}
pvalue=sum(tautestfit$t>=glmm_fit$t)/n_tau
```

```{r echo=FALSE}
print(paste("tau value:",glmm_fit$tau[2]))
print(paste("t value:",glmm_fit$t,"pvalue:",pvalue))
```
```{r permutations, echo=FALSE, fig.height=6, fig.width=8}
ggplot(tautestfit,aes(t))+geom_histogram(bins = 30)+geom_vline(xintercept = glmm_fit$t,color ="red")+theme_minimal(base_size = 16)

```


## Step 3 fit and test beta for each gene
### transform the gene data to a long matrix to fit each gene separately
```{r}
gene_long=exp_genedata %>%
  pivot_longer(cols=starts_with("gene"),
               names_to ="gene_id",
               values_to = "gene_value") %>% 
  as.data.frame()
```

### fit a beta for each gene adjusting for the population structure 
```{r}
gene_test_df=fit_beta(glmm_fit,glm_fit0,GRM,gene_long,SPA=FALSE)
```

### plot a simple volcano plot of the results
```{r, fig.height=6, fig.width=8}
ggplot(gene_test_df,aes(beta,-log10(SPA_pvalue)))+geom_point()+theme_minimal()
```
